How to write a PhD in Biological Sciences: a guide for the uninitiated

### 2021-07-20 3.12-SuppInfo.Rmd
Add files via upload
# After copy edits

## Supplementary Material {#suppinfo}

@@

### What can go into Supplementary Material?
Practically anything!

@@

### There has been a big movement toward transparency and repeatability of scientific studies
This is a really positive movement in science and it should receive all of our support as it is part of the greater scientific project [see part 2](transparency.html). Being proud of your work is being proud of the entire product, including the data and the analysis. If you felt they were good enough for peer review, then you should be happy to make them all freely available. Not sure if you haven’t made some mistakes? We all make mistakes and you shouldn’t think that you will have an entirely fault free career. Whenever and wherever possible conduct your professional due diligence and be open and honest with your research. Science is a career when we should welcome errors and mistakes that we have made in the past. There’s no problem with this. Indeed, the problems only start when you refuse to acknowledge past errors and mistakes, making you prone to motivated reasoning and confirmation bias.

There are lots of data repositories out there. Some specialise in particular kinds of data, like genetic data in [GenBank](https://www.ncbi.nlm.nih.gov/genbank/), or sound data in [The Macaulay Library](https://www.macaulaylibrary.org/) at the Cornell Lab of Ornithology. Others are general repositories of data and you can post anything there, including the scripts that you used to analyse the data. By doing this you ensure that any people that want to use your work in the future have full access to it. Your data deposit will be given a DOI. You can use this DOI in your publication so that readers know where to find the data.

Once I happily allowed the publishers to store my data. These days, I simply don’t trust them to curate and keep my data for posterity. The way that most publishers of academic journals behave, there is no reason for them to make sure that my data is safe and secure. Therefore, I would rather go to an independent curator of data that specialises in not-for-profit data curation. One example is [Zenodo](https://zenodo.org/) which is built and operated by CERN and OpenAIRE. This site has the advantage that it is integrated with GitHub (if you are already a user, but check it out if you aren’t). You can choose to open your data immediately, or embargo it until your paper is published. It’s a great platform, and so much more secure than using a publisher. Another example is the [Open Science Framework (OSF)](https://osf.io/) that aims to be a one stop shop for all of your storage needs. OSF best practices can be found here. You will need to make sure that your data is correctly sorted with sufficient metadata for someone else to understand [see @roche2015public].

Once we did statistical analysis with a GUI and a point and click approach. Forget to click a checkbox and the results were not repeatable. As the analyses have become increasingly sophisticated with more and differing pieces of software, it is becoming increasingly important to record the code that you write that gets your results together with your data. This is the difference between making your results and therefore your study repeatable.

@@

### Why stop with just depositing your data and analyses online?
Let’s face it, you probably put a lot of effort into proposing your thesis hypotheses. Why not deposit this too? Remember that it does not have to be accessible to the world, but it would improve the integrity of your study if your original hypotheses are the same as the ones that you submit for publication. Some journals are now asking whether you have stored hypotheses ahead of conducting the experiments. They don’t stop you from submitting if you haven’t. But why not? It’s a great step forward in terms of transparency and provides more credibility when you eventually write that manuscript with the same aims. And it shouldn’t stop you from publishing another study that happened along the way, even though you hadn’t planned it. I suspect that in future funders will require us to deposit our successful funding proposals in an online repository before they release funds - hopefully they’ll also be willing to support these facilities financially.

>>>>

# Supplementary Material {#suppinfo}

@@

## What can go into Supplementary Material?
Practically anything can go into supplementary material.

@@

## Toward transparency and repeatability of scientific studies
This is a really positive movement in science and it should receive all of our support as it is part of the greater scientific project ([see part 2](#transparency)). Being proud of your work is being proud of the entire product, including the data and the analysis. If you felt they were good enough for peer review, then you should be happy to make them all freely available. Not sure if you haven’t made some mistakes? We all make mistakes and you shouldn’t think that you will have an entirely fault-free career. Whenever and wherever possible conduct your professional due diligence and be open and honest with your research. Science is a career when we should openly acknowledge errors and mistakes that we have made in the past. There’s no problem with this. Indeed, the problems only start when you refuse to acknowledge past errors and mistakes, making you prone to motivated reasoning and confirmation bias.

There are lots of data repositories out there. Some specialise in particular kinds of data, like genetic data in [GenBank](https://www.ncbi.nlm.nih.gov/genbank/), or sound data in [The Macaulay Library](https://www.macaulaylibrary.org/) at the Cornell Lab of Ornithology. Others are general repositories of data and you can post anything there, including the scripts that you used to analyse the data. By doing this you ensure that any people that want to use your work in the future have full access to it. Your data deposit will be given a [DOI](#DOI). You can use this [DOI](#DOI) in your publication so that readers know where to find the data.

Once I happily allowed the publishers to store my data. These days, I simply don’t trust them to curate and keep my data for posterity. The way that most publishers of academic journals behave, there is no reason for them to make sure that my data is safe and secure. Therefore, I would rather go to an independent curator of data that specialises in not-for-profit data curation. One example is [Zenodo](https://zenodo.org/) which is built and operated by CERN and OpenAIRE. This site has the advantage that it is integrated with GitHub (if you are already a user, but check it out if you aren’t). You can choose to open your data immediately, or embargo it until your paper is published. It’s a great platform, and so much more secure than using a publisher. Another example is the [Open Science Framework (OSF)](https://osf.io/) that aims to be a one-stop shop for all of your storage needs. You will need to make sure that your data is correctly sorted with sufficient metadata for someone else to understand [see @roche2015public].

Once, we did statistical analysis with a GUI and a point and click approach. Forget to click a checkbox and the results were not repeatable. As the analyses have become increasingly sophisticated with more and differing pieces of software, it is becoming increasingly important to record the code that you write that gets your results together with your data. This is the difference between making your results and, therefore, your study repeatable.

@@

## Why stop with just depositing your data and analyses online?
Let’s face it, you probably put a lot of effort into proposing your thesis hypotheses. Why not deposit this too? Remember that it does not have to be accessible to the world, but it would improve the integrity of your study if your original hypotheses are the same as the ones that you submit for publication. Some journals are now asking whether you have stored hypotheses ahead of conducting the experiments. They don’t stop you from submitting if you haven’t. But why not? It’s a great step forward in terms of transparency and provides more credibility when you eventually write that manuscript with the same aims. And it shouldn’t stop you from publishing another study that happened along the way, even though you hadn’t planned it. I suspect that in future funders will require us to deposit our successful funding proposals in an online repository before they release funds - hopefully, they’ll also be willing to support these facilities financially.

### 2021-07-20 4.00a-finishing.Rmd
Add files via upload
# After copy edits

# Now that you have finished your data chapters {#part4}

In this last part of the book, I am going to presume that you have done all of the work for the data chapters, and written them up as distinct chapters of your thesis. 
When you look at PhDs that are stored in your institutions libraries, or those that you've referred to in the course of your study, you will notice that they have two additional chapters in addition to data chapters. These are often quite short (although lengths vary), and serve to introduce and conclude the PhD thesis (Figure \@ref(fig:thesis-construction)). While these are not usually considered to be the 'meat' of the PhD itself, they are important sections to reflect on and carefully consider, as they attempt to pull together data chapters that are written as stand-alone documents into a cohesive thesis. To this end, they require the same careful planning that any other section in your thesis deserves, with one important condition: that they should not impede the timely submission of your thesis.

@@

```{r thesis-construction, echo=FALSE, fig.align='left', out.width='90%',  fig.cap="(ref:thesis-construction)"}
knitr::include_graphics('figures/thesis-construction.png')
```

At this point, we have to acknowledge that different institutions require different deadlines of their students in order to finish their PhD studies and submit their theses for examination. In most institutions (to my knowledge) the time to submission is the most important, and can vary between 3 and 5 years (most institutions stipulate a minimum period of 2 years and some have no maximum duration). This period is critical with respect to funding. If funding ceases and your institution allows you to continue your studies, you may have to raise the funding to pay fees as well as to maintain your living costs. After submission there will be a period of time while your thesis is in examination, and another period before the final graduation. Each of these milestones comes with its own set of deadlines and conditions. It is well worth being aware of all these issues well in advance of any submission deadline (ask your advisor). Important aspects for you to note are whether or not you need to remain registered at your institution after you have submitted your thesis, and whether or not your institution will require payment for your registration.

@@

The next chapters deal with writing the [introductory chapter](PhDintro.html), the [concluding chapter](PhDconc.html), and then finally [formatting your thesis for submission](formatting.html). These final hurdles require some time and thought. Given that not all of you will have all of the time that you need, this is often the time when the demand on your writing skills reaches a peak. If you are in this position, I urge you to use your time wisely and to plan carefully what you will write.

>>>>

# Now that you have finished your data chapters

In this last part of the book, I am going to presume that you have done all of the work for the data chapters, and written them up as distinct chapters of your thesis. When you look at PhDs that are stored in your institution's libraries, or those that you’ve referred to in the course of your study, you will notice that they have two additional chapters in addition to data chapters. These are often quite short (although lengths vary), and serve to introduce and conclude the PhD thesis (Figure \@ref(fig:thesis-construction). While these are not usually considered to be the ‘meat’ of the PhD itself, they are important sections to reflect on and carefully consider, as they attempt to pull together data chapters that are written as stand-alone documents into a [cohesive thesis](#criteria). To this end, they require the same careful planning that any other section in your thesis deserves, with one important condition: that they should not impede the timely submission of your thesis.

A typical construction of the data chapters in your thesis together with a introduction and conclusion. The exact construction of your thesis may vary, especially with the number of data chapters that you have. The relationship between thesis chapters may also not be linear as presented here.

@@

```{r thesis-construction, echo=FALSE, fig.show='hold', fig.align='center', out.width='90%', fig.cap="(ref:thesis-construction)"}
knitr::include_graphics('figures/thesis-construction.png')
```

At this point, we have to acknowledge that different institutions require different deadlines of their students in order to finish their PhD studies and submit their theses for examination. In most institutions (to my knowledge) the time to submission is the most important, and can vary between 3 and 5 years (most institutions stipulate a minimum period of 2 years and some have no maximum duration). This period is critical with respect to funding. If funding ceases and your institution allows you to continue your studies, you may have to raise the funding to pay fees as well as to maintain your living costs. After submission, there will be a period of time while your thesis is in examination, and another period before the final graduation. Each of these milestones comes with its own set of deadlines and conditions. It is well worth being aware of all these issues well in advance of any submission deadline (ask your advisor). Important aspects for you to note are whether or not you need to remain registered at your institution after you have submitted your thesis, and whether or not your institution will require payment for your registration.

@@

The next chapters deal with writing the [introductory chapter](#PhDintro), the [concluding chapter](#PhDconc), and then finally [formatting your thesis](#formatting) for submission. These final hurdles require some time and thought. Given that not all of you will have all of the time that you need, this is often the time when the demand on your writing skills reaches a peak. If you are in this position, I urge you to use your time wisely and to plan carefully what you will write.

### 2021-07-20 4.01-PhDintro.Rmd
Add files via upload
# After copy edits

For each of your data chapters, you will have already written an [introduction section](introduction.html), and the same principles that you used to construct this will apply here. Instead of aiming to introduce any one of the hypotheses in the chapters, however, your introduction will aim to introduce the ideas and concepts explored inside the data chapters of the thesis.

My suggestion would be for you to start with the conceptual diagram of the thesis that you produced in your [proposal](proposal.html). If you didn't construct one then, you may be reluctant to do it now, but having a good conceptual idea of how the components of your thesis fit together will really aid you in putting together both this introduction and the final conclusion. For your [proposal](proposal.html), you may also have written a general introduction, and it would be worth revisiting this now, even if you don't actually use any of it.

@@

Next produce a hierarchy of hypotheses for all of the chapters and concepts in your thesis (see figure \@ref(fig:JeschkeHagar61)). This will help you to make sure that you cover all of the [big ideas](bigidea.html) and concepts in the outline. Your outline should follow the standard funnel [formula](formula.html) (see Figure \@ref(fig:introduction)) this time with the highest level ideas in your hierarchy of hypotheses given full prominance in the first paragraph.

@@

The way in which to cite your chapters is most straightfoward here by using their chapter number (i.e. **Chapter n**). This will make it obvious to the examiner (or any other reader) that it is within the content of the thesis. If chapters are published, you could add a citation to the paper in addition, but not instead of.

@@

By the end of the introductory chapter, anyone reading should know what the thesis is about, and what it is that you aim to find out. This doesn't mean that they know what you found in each of the data chapters, but that they know the topics explored within the thesis and what was done to address them. Specifically

@@

You may decide that some of the contents of (particularly data chapter 1) might be more appropriate if it was moved to the introductory chapter. Before you do this I'd suggest that you discuss it with your advisor. Each of the data chapters should have already been signed off (by your advisor) and so it would be a mistake to start chopping them now without consultation. It is also perfectly legitimate to refer in the introduction to a section of the thesis where a particular topic is dealt with in detail. If you are unsure, you'd you best advised by your advisor.

>>>>

For each of your data chapters, you will have already written an [introduction section](#introduction), and the same principles that you used to construct this will apply here. Instead of aiming to introduce any one of the hypotheses in the chapters, however, your introduction will aim to introduce the ideas and concepts explored inside the data chapters of the thesis.

My suggestion would be for you to start with the conceptual diagram of the thesis that you produced in your [proposal](#proposal). If you didn't construct one then, you may be reluctant to do it now, but having a good conceptual idea of how the components of your thesis fit together will really aid you in putting together both this introduction and the final conclusion. For your [proposal](#proposal), you may also have written a general introduction, and it would be worth revisiting this now, even if you don't actually use any of it.

@@

Next, produce a hierarchy of hypotheses for all of the chapters and concepts in your thesis (see Figure \@ref(fig:JeschkeHagar61)). This will help you to make sure that you cover all of the [big ideas](#bigidea) and concepts in the outline. Your outline should follow the standard funnel [formula](#formula) (see Figure \@ref(fig:introduction)) this time with the highest level ideas in your hierarchy of hypotheses given full prominence in the first paragraph.

@@

The way in which to cite your chapters is most straightforward here by using their chapter number (i.e. **Chapter n**). This will make it obvious to the examiner (or any other reader) that it is within the content of the thesis. If chapters are published, you could add a citation to the paper in addition, but not instead of.

@@

By the end of the introductory chapter, anyone reading should know what the thesis is about, and what it is that you aim to find out. This doesn't mean that they know what you found in each of the data chapters, but that they know the topics explored within the thesis and what was done to address them.

@@

You may decide that some of the contents of (particularly data chapter 1) might be more appropriate if it was moved to the introductory chapter. Before you do this I'd suggest that you discuss it with your advisor. Each of the data chapters should have already been signed off (by your advisor) and so it would be a mistake to start chopping them now without consultation. It is also perfectly legitimate to refer in the introduction to a section of the thesis where a particular topic is dealt with in detail. If you are unsure, you'd best be advised by your advisor.

### 2021-07-20 4.02-PhDconc.Rmd
Add files via upload
# After copy edits

By the time you come to write your concluding chapter of your PhD, you should have written all of the [data chapters](part3.html) and the [introductory chapter](PhDintro.html). Firstly, I should congratulate you for getting this far. You've almost finished, and there's just one final push required to complete the document. You are now the best placed person to tell yourself what should be in the concluding chapter.

## Using the funnel structure
In principle, you can use the same funnel structure that we described for the [discussion](discussion.html) as a way of structuring your concluding chapter (see figure \@ref(fig:discussion)). This would include starting with a synopsis of what your thesis is trying to achieve. Clearly, this starting section may be more than a paragraph (as it was in the [data chapters](part3.html)), and the precise structure will be dictated by the complexity of your thesis structure (Figure \@ref(fig:thesis-construction)).

@@

I suggest that you take your conceptual outline (from the [introductory chapter](PhDintro.html)), and use this to being thinking about how best to construct the outline of your concluding chapter. If your data chapters are used to feed into a final output, then consider using the approach of a [logical argument](argument.html) as the outline.

The main body should review the concluding findings of each of the data chapters, and especially the way in whcih they interact - as these are the aspects that are least likely to have been touched on already in the individual data chapters. It is hoped that the thesis represents a body of work, and not a series of unconnected works. Thus your task in this final chapter is to present to the examiner (and any other reader) exactly how the chapters fit together into a body of work.

Your concluding chapter should provide all of the highlights and insights that have been gained from your PhD thesis. It is unlikely to you need to dwell on individuals caveats from the chapters, unless there are general caveats from your approach that span chapters - in which case these probably deserve attention.

@@

You also have an opportunity to point to where, in your opionion, the direction of your particular field should move next. **Do provide your opinion**. Remember that you are now one of the most qualified people in this particular area to be pronouncing on what should be done. Your opinion is therefore valid, and your concluding chapter is a good place to provide a reasoned argument about your own ideas.

>>>>

By the time you come to write your concluding chapter of your PhD, you should have written all of the [data chapters](#part3) and the [introductory chapter](#PhDintro). Firstly, I should congratulate you for getting this far. You've almost finished, and there's just one final push required to complete the document. You are now the best placed person to tell yourself what should be in the concluding chapter.

## Using the funnel structure
In principle, you can use the same funnel structure that we described for the [discussion](#discussion) as a way of structuring your concluding chapter (see Figure \@ref(fig:discussion)). This would include starting with a synopsis of what your thesis is trying to achieve. Clearly, this starting section may be more than a paragraph (as it was in the [data chapters](#part3)), and the precise structure will be dictated by the complexity of your thesis structure (Figure \@ref(fig:thesis-construction)).

@@

I suggest that you take your conceptual outline (from the [introductory chapter](#PhDintro)), and use this to being thinking about how best to construct the outline of your concluding chapter. If your data chapters are used to feed into a final output, then consider using the approach of a [logical argument](#argument) as the outline.

The main body should review the concluding findings of each of the data chapters, and especially the way in which they interact - as these are the aspects that are least likely to have been touched on already in the individual data chapters. It is hoped that the thesis represents a body of work, and not a series of unconnected works. Thus your task in this final chapter is to present to the examiner (and any other reader) exactly how the chapters fit together into a body of work.

Your concluding chapter should provide all of the highlights and insights that have been gained from your PhD thesis. It is unlikely that you need to dwell on individuals caveats from the chapters, unless there are general caveats from your approach that span chapters - in which case, these probably deserve attention.

@@

You also have an opportunity to point to where, in your opinion, the direction of your particular field should move next. **Do provide your opinion**. Remember that you are now one of the most qualified people in this particular area to be pronouncing on what should be done. Your opinion is therefore valid, and your concluding chapter is a good place to provide a reasoned argument about your own ideas.

### 2021-07-20 4.03-formatting.Rmd
Add files via upload
# After copy edits

If you chose [Bookdown](bookdown.org) or [thesisdown](thesisdown.org) to write your thesis in, then you'll simply be able to download the correct style file (cls) and bibliography formatting file (bib) and press the build book button in [RStudio](www.rstudio.com) and you'll have your formatted thesis in the time that it takes your computer to render the pdf. You'll still need to look through it and make sure that the output is what you want.

@@

Be patient and don't try to build mega-big files in word-processing documents, they tend to crash. Instead, copy the template file and compile one chapter at a time, placing them together only at the end for a final build.

Your university will have their own stipulation for exactly how they want the thesis formatted. Here, I provide a check-list for the last things to do before you submit. The chances are that your university will have a bunch of helpful links. Make sure that you follow the guidelines of your own institution well in advance of submitting your thesis.

### Must do check-list before submitting your thesis:
The following points are based on what may irritate or annoy your examiner (and you really don’t want to do that!):

@@

- Make sure that your language settings are set to ‘UK English’ (or the English setting for where you are based. 

@@

- Capitalisation of [common names](taxonomy.html), place names and not adjectives. For example, ‘South Africa’ has two capitals, but ‘southern Africa’ only has one.

@@

- Pay attention if you have used “we” or “I” and make sure they are consistent in your thesis. As a rule I encourage “I” in a thesis, unless the chapter is also a manuscript, in which case “we” is usually correct.

@@

- Sections and subheadings. I’ve encouraged you to use subheadings throughout your thesis. Here you have a chance to number them sequentially. This is very useful for your examiners and may be a requirement for the university. Using the [word processor’s](software.html) built in functions will make this task consistent and easy.

@@

- Remember to give them a check through before handing it in. If you’ve done the sections correctly, then the contents page will come out right.
- [Title page](title:html) - prescribed very strictly by the university. The librarians place the SU watermark after final submission to the library. It may be tempting to change your title now that you know what's in the thesis, but many universities have strict policies about this. If you want to change the title, make sure that you are able to do so.

@@

- [Acknowledgements](acknowledgements.html) - this is your time to say thank you to all the special people that have helped during your study. There are probably more than you realise, but in addition to your friends and family (who most people don’t forget), think about the people who administered the work, lab mates (past and present) who were always there to help, and people who gave permission at study sites. 
- [References](references.html) - probably one of the most dreaded sections of any thesis preparation, but they do have to be done. If you’re one of these people that has everything in a [database](database.html), then you’ll be laughing or cursing your database throughout. While it might be tempting to only look through the data within the database, spend some time to see how it’s displayed in the thesis. A mistake in the reference database will be multiplied many times in the thesis. Remember that examiners love to take a random look through the reference section to make sure that it’s all good. After years of painfully entering references themselves, they know just what to look for.

### Mistakes people make:
Other than the obvious things, all mentioned above, here are some of the mistakes I’ve seen.

- Submitting the wrong version (yes this does happen!). Probably worse if having a mixture of right and wrong versions for different chapters (worse because it takes longer to sort out)
- Last minute additions to text with incorrect spelling and or grammar
- Two correctly spelled synonyms sitting next to each other when only one is desired (probably came about when editing)
- Forgetting to check for plagiarism (see here)
- Comments and or edited text (especially when it’s marked as being by someone other than the student).
- Page numbers that start again and again at different sections
- Lots of blank pages or spaces (avoid blank pages if you can, and try to limit the amount of blank space (never >half a page).
- Leaving important people out of the acknowledgements (e.g. advisor, administrators, funders, etc.)

>>>>

If you chose [Bookdown](bookdown.org) or [Thesisdown](thesisdown.org) to write your thesis in, then you'll simply be able to download the correct style file (cls) and bibliography formatting file (bib) and press the build book button in [RStudio](www.rstudio.com) and you'll have your formatted thesis in the time that it takes your computer to render the pdf. You'll still need to look through it and make sure that the output is what you want.

@@

Be patient and don't try to build mega-big files in word-processing documents, as they tend to crash. Instead, copy the template file and compile one chapter at a time, placing them together only at the end for a final build.

Your university will have its own stipulation for exactly how they want the thesis formatted. Here, I provide a check-list for the last things to do before you submit. The chances are that your university will have a bunch of helpful links. Make sure that you follow the guidelines of your own institution well in advance of submitting your thesis.

## Must do check-list before submitting your thesis:
The following points are based on what may irritate or annoy your examiner (and you really don’t want to do that!):

@@

- Make sure that your language settings are set to ‘UK English’ (or the English setting for where you are based). 

@@

- Capitalisation of [common names](#taxonomy), place names and not adjectives. For example, ‘South Africa’ has two capitals, but ‘southern Africa’ only has one.

@@

- Pay attention if you have used 'we' or 'I' and make sure they are consistent in your thesis. As a rule, I encourage 'I' in a thesis, unless the chapter is also a manuscript, in which case 'we' is correct.

@@

- Sections and subheadings. I’ve encouraged you to use subheadings throughout your thesis. Here you have a chance to number them sequentially. This is very useful for your examiners and may be a requirement for the university. Using the [word processor’s](#software) built in functions will make this task consistent and easy.

@@

- Remember to give them a check through before handing it in. If you’ve done the sections correctly, then the contents page will come out correctly.
- [Title page](#title) - prescribed very strictly by the university. The librarians place a watermark after final submission to the library. It may be tempting to change your title now that you know what's in the thesis, but many universities have strict policies about this. If you want to change the title, make sure that you are able to do so.

@@

- [Acknowledgements](#acknowledgements) - this is your time to say thank you to all the special people that have helped during your study. There are probably more than you realise, but in addition to your friends and family (who most people don’t forget), think about the people who administered the work, lab mates (past and present) who were always there to help, and people who gave permission at study sites. 
- Check that you provide links to any [online repositories](#suppinfo) of your data.
- [References](#references) - probably one of the most dreaded sections of any thesis preparation, but they do have to be done. If you’re one of these people that has everything in a [database](#database), then you’ll be laughing or cursing your database throughout. While it might be tempting to only look through the data within the database, spend some time to see how it’s displayed in the thesis. A mistake in the reference database will be multiplied many times in the thesis. Remember that examiners love to take a random look through the reference section to make sure that it’s all good. After years of painfully entering references themselves, they know just what to look for.

## Mistakes people make:
Other than the obvious things, all mentioned above, here are some of the mistakes I’ve seen.

- Submitting the wrong version (yes, this does happen!). Probably worse if having a mixture of right and wrong versions for different chapters (worse because it takes longer to sort out).
- Last-minute additions to text with incorrect spelling and or grammar.
- Two correctly spelled synonyms sitting next to each other when only one is desired (probably came about when editing).
- Forgetting to check for [plagiarism](#plagiarism).
- Comments and or edited text (especially when it’s marked as being by someone other than the student).
- Page numbers that start again and again at different sections.
- Lots of blank pages or spaces (avoid blank pages if you can, and try to limit the amount of blank space (never >half a page).
- Leaving important people out of the acknowledgements (e.g. advisor, administrators, funders, etc.).

### 2021-07-20 4.03-publish.Rmd
Add files via upload
# After copy edits

# The obligatoin to publish your work {.unnumbered #publish}

Now that you've submitted your PhD, you need still need to do something with it. For most of you this will be publishing it in a journal of your choice. It will then take you through what happens to the manuscript and give you insight about where you might want to publish, how to choose the journal. Publication is important. Here is a list of reasons why you have some obligation to publish the work in your thesis:

- If your work is unpublished, then it will not be used. Without use, all the work that you put into it falls away.
- Your published work will become the foundation for future students and academics who are active in your field. This might well be in another country or on another continent.
- Your work was most likely paid for by tax-payers in the country where you studied. Publishing this work is a way of handing back the value of what you found. If it is published Open Access, the public can also read it for themselves. 
- Other funders of your thesis work may have made publication a prerequisite of the funding criteria.
- It is hoped that by publishing your work it will become more accessible to the scientific community.
- By publishing your work you will find that both you, and your work, become known by an international community that may well invite you to participate in the academic process further [@marks2013misuse].
- Both inside and outside of science as a career, peer reviewed published papers are seen as an important accomplishment in addition to the production of your thesis.
- The work that you have already put into your thesis (especially if you have followed the advice in this book!), will mean that you are close to having chapters that can be submitted as publications. 
- Your advisor and lab colleagues who helped with your thesis work may depend on your publications. 
- There is a lot of satisfaction to be had by seeing your work published. It has been argued that this depends on who you are and where you come from [@husemann2017publicationism]. 
- The satisfaction of having someone say how enjoyable or inspiring your work is to read will certainly not be as frequent as if you managed to publish a novel or newspaper column. But when it does happen, it will bring a smile to your face.

This is by no means an exhaustive list, but I hope it will give you some insight into the importance of publishing your results. To counter this assertion, please be aware that there is also research that suggests that publication driven science is not healthy, and that we need a new way to motivate science [e.g. @stergiou2014impact]. Indeed, it has been suggested that the ‘publish or perish’ mentality of academia has resulted in the retention (and even deterioration) of poor experimental design, and hence higher false positives in the behavioural sciences [@smaldino2016natural]. However, before you can join the debate about the future, I’d argue that you need to provide your credentials for the present.

Taking your work to the next level is not a trivial step, even though you already have your chapters written in the style of manuscripts. The challenges of publishing work, like other challenges in academia, are much easier if you possess inside knowledge of the system. The demystification of the next steps along the road in academia are the subject of another book:

>>>>

# The obligation to publish your work {.unnumbered #publish}

Now that you've submitted your PhD, you still need need to do something with it. For most of you this will be publishing it in a journal of your choice. Publication is important. Here is a list of reasons why you have some obligation to publish the work in your thesis:

- If your work is unpublished, then it will not be used. Without use, all the work that you put into it falls away.
- Your published work will become the foundation for future students and academics who are active in your field. This might well be in another country or on another continent.
- Your work was most likely paid for by tax-payers in the country where you studied. Publishing this work is a way of handing back the value of what you found. If it is published Open Access, the public can also read it for themselves. 
- Other funders of your thesis work may have made publication a prerequisite of the funding criteria.
- It is hoped that by publishing your work it will become more accessible to the scientific community.
- By publishing your work you will find that both you, and your work, become known by an international community that may well invite you to participate in the academic process further [@marks2013misuse].
- Both inside and outside of science as a career, peer-reviewed published papers are seen as an important accomplishment in addition to the production of your thesis.
- The work that you have already put into your thesis (especially if you have followed the advice in this book!), will mean that you are close to having chapters that can be submitted as publications. 
- Your advisor and lab colleagues who helped with your thesis work may depend on your publications. 
- There is a lot of satisfaction to be had by seeing your work published. It has been argued that this depends on who you are and where you come from [@husemann2017publicationism]. 
- The satisfaction of having someone say how enjoyable or inspiring your work is to read will certainly not be as frequent as if you managed to publish a novel or newspaper column. But when it does happen, it will bring a smile to your face.

This is by no means an exhaustive list, but I hope it will give you some insight into the importance of publishing your results. To counter this assertion, please be aware that there is also research that suggests that publication driven science is not healthy, and that we need a new paradigm to motivate science [e.g. @stergiou2014impact]. Indeed, it has been suggested that the ‘publish or perish’ mentality of academia has resulted in the retention (and even deterioration) of poor experimental design, and hence higher false positives in the behavioural sciences [@smaldino2016natural]. However, before you can join the debate about the future, I’d argue that you need to provide your credentials for the present.

Taking your work to the next level is not a trivial step, even though you already have your chapters written in the style of manuscripts. The challenges of publishing work, like other challenges in academia, are much easier if you possess inside knowledge of the system. The demystification of the next steps along the road in academia is the subject of another book.

```{r cover2, echo=FALSE, fig.align='centre', out.width='60%', fig.link='www.routledge.com/9781032116419'}
knitr::include_graphics('figures/cover2.png')
```

**The follow-up to this book, How to Publish in Biological Sciences: A Guide for the Uninitiated**  is available at [www.routledge.com/9781032116419](www.routledge.com/9781032116419).

### 2021-07-20 4.04-lastnote.Rmd
Add files via upload
# After copy edits

If you have found this book useful, then please remember that you can contribute to the project at the Github pages for this book, using bookdown. If your contributions, no matter how small, improve the book, then you will be improving the experience of doing a PhD in Biological Sciences for future PhD students, and that has to be worth something.

>>>>

If you have found this book useful, then please remember that you can [contribute to the project](#contribute) at the Github pages for this book, using bookdown. If your contributions, no matter how small, improve the book, then you will be improving the experience of doing a PhD in Biological Sciences for future PhD students, and that has to be worth something.
